#### DEA {#resultsRnaseqSRP084288DEA}

##### Fluoxetine VS Control with Tissue as covariate{#resultsRnaseqSRP084288FluoxetineVsControlWithTissueAsCovariate}

###### Apeglm shrinkage {#resultsRnaseqSRP084288FluoxetineVsControlWithTissueAsCovariateApeglm}

```{r resultsRnaseqSRP084288FluoxetineVsControlWithTissueAsCovariatesApeglmComputeDeseq2}
dds <- DESeqDataSetFromMatrix(
  countData = SRP084288_raw[, -1],
  colData = SRP084288_metadata,
  rowData = SRP084288_genes,
  design = ~ tissue_region + group
)

# Unsure to filter genes here
# Because I want to be able to see them
# if they are DE in another experiment.
keep <- rowSums(counts(dds)) >= ncol(counts(dds))
dds <- dds[keep, ]

dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

# vsd <- varianceStabilizingTransformation(dds)

res <- lfcShrink(
  dds,
  coef = "group_Fluoxetine_vs_Control",
  type = "apeglm"
)
```

```{r resultsRnaseqSRP084288FluoxetineVsControlWithTissueAsCovariateApeglmKnitPostDeseq2, results="asis", cache=FALSE}

knitPostDeseq2Template(
  FC_cutoff = 0.5,
  cols = rat_cols
)

res$project <- "SRP084288"
res$comparison <- "FluoxetineVsControlWithTissueAsCovariate"

dea_res$rnaseq$SRP084288$FluoxetineVsControlWithTissueAsCovariate <- cbind(
  mcols(dds)[, 1:4],
  res
)
```


```{r resultsRnaseqSRP084288FluoxetineVsControlWithTissueAsCovariatesApeglmDefaultSettingsHeatmapSettings}
fig_height <- 30
n_top_genes <- 200

dds <- varianceStabilizingTransformation(dds)
dds <- dds[rank(res$pvalue) < 200, ]
```

```{r resultsRnaseqSRP084288FluoxetineVsControlWithTissueAsCovariatesApeglmDefaultSettingsHeatmapRaw, dependson="resultsRnaseqSRP084288FluoxetineVsControlWithTissueAsCovariatesDefaultSettingsHeatmapSettings", fig.height=fig_height}
set.seed(123)
Heatmap(
  assay(dds),
  row_labels = rowData(dds)$hugo_symbol,
  top_annotation = HeatmapAnnotation(
    df = colData(dds)[
      ,
      c(
        "tissue_region",
        "group"
      )
    ]
  ),
  name = "RNA signal",
  column_title = "Samples",
  row_title = "Genes"
)
```

```{r resultsRnaseqSRP084288FluoxetineVsControlWithTissueAsCovariatesApeglmDefaultSettingsHeatmapZscore, dependson="resultsRnaseqSRP084288FluoxetineVsControlWithTissueAsCovariatesDefaultSettingsHeatmapSettings", fig.height=fig_height}
set.seed(123)
Heatmap(
  scaleRows(assay(dds)),
  row_labels = rowData(dds)$hugo_symbol,
  top_annotation = HeatmapAnnotation(
    df = colData(dds)[
      ,
      c(
        "tissue_region",
        "group"
      )
    ]
  ),
  name = "RNA signal",
  column_title = "Samples",
  row_title = "Genes"
)
```

This heatmap was used to check fold change sign is correct for this design:
We can see ASAP1 is overexpressed with Fluoxetine, which can be consistently seen on integration dotplots for the design `GSE84185-FluoxetineVsControlForNoStressedInDentateGyrus`.
