#### Rat+Mouse Fluoxetine VS Control for Stressed in Dentate Gyrus{#resultsIntegrationOriginalSpecieRatMouseFluoxetineVsControlForStressed}

This section corresponds to the first+second meta-analysis requested by El Chérif.

```{r resultsIntegrationOriginalSpecieRatMouseFluoxetineVsControlForStressedSettings}
fig_width <- 42
fig_height <- 32
n_cols <- 17
```

```{r resultsIntegrationOriginalSpecieRatMouseFluoxetineVsControlForStressedAlphabet, dependson="resultsIntegrationOriginalSpecieRatMouseFluoxetineVsControlForStressedSettings", fig.height=fig_height, fig.width=fig_width, fig.cap="Alphabetically-sorted genes significant in at least one comparison (padj < 0.05). + and – signs highlight respectively upregulated and downregulated genes (|logFoldChange| > 0.1 and raw pvalue < 0.05)."}
cols_to_keep <- c(
  "Gene.name",
  "baseMean",
  "log2FoldChange",
  "pvalue",
  "padj",
  "project",
  "comparison"
)

# First two columns are removed because they are specie specific
d <- data.table(
  rbind(
    data.frame(dea_res$microarray_v2$GSE43261$FluoxetineResponderVsControl[, cols_to_keep]),
    #data.frame(dea_res$microarray_v2$GSE84185$StressedFluoxetineVsStressedAloneInDentateGyrus[, cols_to_keep]),
    data.frame(dea_res$microarray_v2$GSE84185$FluoxetineVsControlForStressedResponderInDentateGyrus[, cols_to_keep]),
    data.frame(dea_res$microarray_v2$Surget2009$FluoxetineVsSalineForStressedInDentateGyrus[, cols_to_keep]),
    data.frame(dea_res$rnaseq$SRP057486$FluoxetineVsVehicleForBrainCA1StressedSubset[, cols_to_keep]),
    data.frame(dea_res$microarray_v2$GSE56028$FluoxetineVsUntreatedForStressed[, cols_to_keep]),
    data.frame(dea_res$rnaseq$SRP131063$FluoxetineVsSalineForStressed[, cols_to_keep]),
    data.frame(dea_res$rnaseq$SRP084288$FluoxetineVsModelOfDepressionWithTissueAsCovariate[, cols_to_keep])
  )
)
d <- d[!is.na(d$pvalue), ]
#d <- d[!d$Gene.name %in% "N/A", ]
d$design <- paste(
  d$project,
  d$comparison,
  sep = "-"
)
designs <- factor(sort(unique(d$design)))
design_ids <- paste(
  as.integer(designs),
  sub(
    "-.*$",
    "",
    designs
  ),
  sep = "-"
)
d$design <- mapvalues(
  d$design,
  from = designs,
  to = design_ids
)

knitr::kable(
  data.frame(
    design_ids,
    designs
  ),
  caption = "Mapping between ids and designs used in the integration plots below"
)

d <- d[order(pvalue)]
d[, duplicated_Gene_name := duplicated(Gene.name), , by = design]
d <- d[!(duplicated_Gene_name)]
d[, pval_rank := rank(pvalue), by=design]
# Add columns to compute portrait score like Stephen Gammie 2019
d[, score := -log10(pvalue) * sign(log2FoldChange)]
d[, score_up_rank := rank(-score), by = design]
d[, score_dn_rank := rank(score), by = design]

genes <- unique(d$Gene.name)
names(genes) <- genes
vec_portrait_scores <- sapply(
  genes,
  computeGammie2019PortraitScore
)
vec_consensus_scores <- sapply(
  genes,
  computeConsensusScore
)
d_scores <- data.frame(
  gene = names(vec_consensus_scores),
  consensus_score = vec_consensus_scores,
  portrait_score = vec_portrait_scores
)
d_scores <- d_scores[
  order(
    d_scores$consensus_score + d_scores$portrait_score,
    decreasing = TRUE
  ),
]
rm(
  vec_consensus_scores,
  vec_portrait_scores
)

d <- data.frame(d)
# Working on a subset because not all genes have a padj from Deseq2
#ds <- d[!is.na(d$padj), ]
genes_to_display <- unique(
  sort(
    d[
      d$pvalue < 0.05 &
      d$pval_rank < 301,
      "Gene.name"
    ]
  )
)

pl <- list() # plot list
gl <- list() # gene list
for (n in 1:n_cols) {
  first_idx <- ceiling((n - 1) / n_cols * length(genes_to_display)) + 1
  last_idx <- ceiling(n / n_cols * length(genes_to_display))
  gl[[n]] <- genes_to_display[first_idx:last_idx]
  pl[[n]] <- plotIntegrationDeaDots(
    d[d$Gene.name %in% gl[[n]], ],
    y_col = "Gene.name"
  )
}

ggarrange(
  plotlist = pl,
  ncol = length(pl),
  nrow = 1,
  align = "hv",
  common.legend = TRUE
)
```

The xlsx table with all genes is available [here](xlsx/resultsIntegrationOriginalSpecieRatMouseFluoxetineVsControlForStressedExportAllGenes.xlsx)

```{r resultsIntegrationOriginalSpecieRatMouseFluoxetineVsControlForStressedExportAllGenes}
# Export to xlsx
absMax <- function(x){return(x[which.max(abs(x))])}
#d <- d[!is.na(d$Gene.name),]
d_log2FoldChange <- reshape2::dcast(
  data = data.frame(d),
  formula = Gene.name ~ design,
  value.var = "log2FoldChange",
  fun.aggregate = absMax,
  fill = 0
)
d_log2FoldChange <- d_log2FoldChange[
  match(
    d_scores$gene,
    d_log2FoldChange$Gene.name
  ),
]
d_pvalue <- reshape2::dcast(
  data = d,
  formula = Gene.name ~ design,
  value.var = "pvalue",
  fun.aggregate = min,
  fill = 1
)
d_pvalue <- d_pvalue[
  match(
    d_scores$gene,
    d_pvalue$Gene.name
  ),
]
d_pval_rank <- reshape2::dcast(
  data = d,
  formula = Gene.name ~ design,
  value.var = "pval_rank",
  fun.aggregate = min,
  fill = 99999
)
d_pval_rank <- d_pval_rank[
  match(
    d_scores$gene,
    d_pval_rank$Gene.name
  ),
]
wb <- createWorkbook()
addWorksheet(
  wb,
  "log2FoldChange"
)
addWorksheet(
  wb,
  "pvalue"
)
addWorksheet(
  wb,
  "pval_rank"
)
addWorksheet(
  wb,
  "scores"
)
writeDataTable(
  wb,
  "log2FoldChange",
  d_log2FoldChange,
  startCol = 1
)
writeDataTable(
  wb,
  "pvalue",
  d_pvalue,
  startCol = 1
)
writeDataTable(
  wb,
  "pval_rank",
  d_pval_rank,
  startCol = 1
)
writeDataTable(
  wb,
  "scores",
  d_scores,
  startCol = 1
)
conditionalFormatting(
  wb,
  sheet = "scores",
  rows = 1:nrow(d_scores) + 1,
  type = "colourScale",
  style = c("blue", "white", "red"),
  rule = c(
    -length(designs),
    0,
    length(designs)
  ),
  cols = 2:ncol(d_scores)
)
conditionalFormatting(
  wb,
  sheet = "pvalue",
  rows = 1:nrow(d_pvalue) + 1,
  type = "colourScale",
  style = c("darkgreen", "white"),
  rule = c(0, 0.1),
  cols = 2:ncol(d_pvalue)
)
conditionalFormatting(
  wb,
  sheet = "pval_rank",
  rows = 1:nrow(d_pvalue) + 1,
  type = "colourScale",
  style = c("darkgreen", "white"),
  rule = c(1, 300),
  cols = 2:ncol(d_pvalue)
)
conditionalFormatting(
  wb,
  sheet = "log2FoldChange",
  rows = 1:nrow(d_log2FoldChange) + 1,
  type = "colourScale",
  style = c("blue", "white", "red"),
  rule = c(-3, 0, 3),
  cols = 2:ncol(d_log2FoldChange)
)
for (
  worksheet in c(
    "pvalue",
    "log2FoldChange",
    "pval_rank"
  )
) {
  freezePane(
    wb,
    sheet = worksheet,
    firstRow = TRUE,
    firstCol = TRUE
  )
  headerStyle <- createStyle(
    # fontSize = 18,
    # fontName = "Arial",
    # textDecoration = "bold",
    # halign = "left",
    # fgFill = "#1A33CC",
    # border = "TopBottomLeftRight",
    textRotation = 90
  )
  addStyle(
    wb,
    sheet = worksheet,
    style = headerStyle,
    rows = 1,
    cols = 1:ncol(d_pvalue)
  )
  setRowHeights(
    wb,
    sheet = worksheet,
    rows = 1,
    heights = 200
  )
  setColWidths(
    wb,
    sheet = worksheet,
    cols = 2:ncol(d_pvalue),
    widths = 4,
  )
  setColWidths(
    wb,
    sheet = worksheet,
    cols = 1,
    widths = 8,
  )
  saveWorkbook(
    wb,
    file = file.path(
      book_from_rmd,
      "xlsx",
      paste0(
        opts_current$get("label"),
        ".xlsx"
      )
    ),
    overwrite = TRUE
  )
}
```

```{r resultsIntegrationOriginalSpecieRatMouseFluoxetineVsControlForStressedConsensus, dependson="resultsIntegrationOriginalSpecieRatMouseFluoxetineVsControlForStressedConsensusSettings", fig.height=fig_height, fig.width=fig_width, fig.cap="Consensus-sorted genes significant in at least one comparison (p-value < 0.05 and p-value rank < 301). + and – signs highlight respectively upregulated and downregulated genes (|logFoldChange| > 0.1 and raw pvalue < 0.05). Underlying table is available [here](xlsx/resultsIntegrationOriginalSpecieRatMouseFluoxetineVsControlForStressedConsensus.xlsx)"}
consensus_scores <- sapply(
  genes_to_display,
  computeConsensusScore,
  df = d
)
genes_to_display <- names(
  sort(
    consensus_scores
  )
)
d$Gene.name <- factor(
  d$Gene.name,
  levels = genes_to_display
)
# Export to xlsx
absMax <- function(x){return(x[which.max(abs(x))])}
d <- d[!is.na(d$Gene.name),]
d_log2FoldChange <- reshape2::dcast(
  data = data.frame(d),
  formula = Gene.name ~ design,
  value.var = "log2FoldChange",
  fun.aggregate = absMax,
  fill = 0
)
d_pvalue <- reshape2::dcast(
  data = d,
  formula = Gene.name ~ design,
  value.var = "pvalue",
  fun.aggregate = min,
  fill = 1
)
d_pval_rank <- reshape2::dcast(
  data = d,
  formula = Gene.name ~ design,
  value.var = "pval_rank",
  fun.aggregate = min,
  fill = 99999
)
d_scores <- d_scores[
  match(
    d_pval_rank$Gene.name,
    d_scores$gene
  ),
]
wb <- createWorkbook()
addWorksheet(
  wb,
  "log2FoldChange"
)
addWorksheet(
  wb,
  "pvalue"
)
addWorksheet(
  wb,
  "pval_rank"
)
addWorksheet(
  wb,
  "scores"
)
writeDataTable(
  wb,
  "log2FoldChange",
  d_log2FoldChange,
  startCol = 1
)
writeDataTable(
  wb,
  "pvalue",
  d_pvalue,
  startCol = 1
)
writeDataTable(
  wb,
  "pval_rank",
  d_pval_rank,
  startCol = 1
)
writeDataTable(
  wb,
  "scores",
  d_scores,
  startCol = 1
)
conditionalFormatting(
  wb,
  sheet = "scores",
  rows = 1:nrow(d_scores) + 1,
  type = "colourScale",
  style = c("blue", "white", "red"),
  rule = c(
    -length(designs),
    0,
    length(designs)
  ),
  cols = 2:ncol(d_scores)
)
conditionalFormatting(
  wb,
  sheet = "pvalue",
  rows = 1:nrow(d_pvalue) + 1,
  type = "colourScale",
  style = c("darkgreen", "white"),
  rule = c(0, 0.1),
  cols = 2:ncol(d_pvalue)
)
conditionalFormatting(
  wb,
  sheet = "pval_rank",
  rows = 1:nrow(d_pvalue) + 1,
  type = "colourScale",
  style = c("darkgreen", "white"),
  rule = c(1, 300),
  cols = 2:ncol(d_pvalue)
)
conditionalFormatting(
  wb,
  sheet = "log2FoldChange",
  rows = 1:nrow(d_log2FoldChange) + 1,
  type = "colourScale",
  style = c("blue", "white", "red"),
  rule = c(-3, 0, 3),
  cols = 2:ncol(d_log2FoldChange)
)
for (
  worksheet in c(
    "pvalue",
    "log2FoldChange",
    "pval_rank"
  )
) {
  freezePane(
    wb,
    sheet = worksheet,
    firstRow = TRUE,
    firstCol = TRUE
  )
  headerStyle <- createStyle(
    # fontSize = 18,
    # fontName = "Arial",
    # textDecoration = "bold",
    # halign = "left",
    # fgFill = "#1A33CC",
    # border = "TopBottomLeftRight",
    textRotation = 90
  )
  addStyle(
    wb,
    sheet = worksheet,
    style = headerStyle,
    rows = 1,
    cols = 1:ncol(d_pvalue)
  )
  setRowHeights(
    wb,
    sheet = worksheet,
    rows = 1,
    heights = 200
  )
  setColWidths(
    wb,
    sheet = worksheet,
    cols = 2:ncol(d_pvalue),
    widths = 4,
  )
  setColWidths(
    wb,
    sheet = worksheet,
    cols = 1,
    widths = 8,
  )
  saveWorkbook(
    wb,
    file = file.path(
      book_from_rmd,
      "xlsx",
      paste0(
        opts_current$get("label"),
        ".xlsx"
      )
    ),
    overwrite = TRUE
  )
}
pl <- list() # plot list
gl <- list() # gene list
for (n in 1:n_cols) {
  first_idx <- ceiling((n - 1) / n_cols * length(genes_to_display)) + 1
  last_idx <- ceiling(n / n_cols * length(genes_to_display))
  gl[[n]] <- genes_to_display[first_idx:last_idx]
  pl[[n]] <- plotIntegrationDeaDots(
    d[d$Gene.name %in% gl[[n]], ],
    y_col = "Gene.name"
  )
}
ggarrange(
  plotlist = pl,
  ncol = length(pl),
  nrow = 1,
  align = "hv",
  common.legend = TRUE
)
```
