#### Fluoxetine VS Control For  Stressed in Dentate Gyrus{#resultsMicroarrayGSE84185StressedFluoxetineVsStressedAloneInDentateGyrus}

##### Default settings {#resultsMicroarrayGSE84185StressedFluoxetineVsStressedAloneInDentateGyrusDefaultSettings}

```{r resultsMicroarrayGSE84185StressedFluoxetineVsStressedAloneInDentateGyrusDefaultSettings}
# Version info: R 3.2.3, Biobase 2.30.0, GEOquery 2.40.0, limma 3.26.8
################################################################
#   Differential expression analysis with limma
library(GEOquery)
library(limma)
library(umap)

# load series and platform data from GEO

gset <- getGEO(
  "GSE84185",
  destdir = geo_dir,
  GSEMatrix = TRUE,
  AnnotGPL = FALSE
)

fData(gset[[1]]) <- cbind(
  fData(gset[[1]]),
  annotations$AGILENT_SurePrint_G3_GE_8x60k[
    match(
      fData(gset[[1]])$NAME,
      annotations$AGILENT_SurePrint_G3_GE_8x60k$probe
    )
  ]
)


if (length(gset) > 1) idx <- grep("GPL13912", attr(gset, "names")) else idx <- 1
gset <- gset[[idx]]

# make proper column names to match toptable
fvarLabels(gset) <- make.names(fvarLabels(gset))

# group membership for all samples
gsms <- paste0(
  "XXXXXXXX1111111100000000XXXXXXXXXXXXXXXXXXXXXXXXXX",
  "XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX"
)
sml <- strsplit(gsms, split = "")[[1]]

# filter out excluded samples (marked as "X")
sel <- which(sml != "X")
sml <- sml[sel]
gset <- gset[, sel]

# log2 transformation
ex <- exprs(gset)
qx <- as.numeric(quantile(ex, c(0., 0.25, 0.5, 0.75, 0.99, 1.0), na.rm = T))
LogC <- (qx[5] > 100) ||
  (qx[6] - qx[1] > 50 && qx[2] > 0)
if (LogC) {
  ex[which(ex <= 0)] <- NaN
  exprs(gset) <- log2(ex)
}

# assign samples to groups and set up design matrix
gs <- factor(sml)
groups <- make.names(
  c(
    "Stress_Fluoxetine_Dentate_Gyrus",
    "Stress_Alone_Dentate_Gyrus"
  )
)
levels(gs) <- groups
gset$group <- gs
design <- model.matrix(~ group + 0, gset)
colnames(design) <- levels(gs)

fit <- lmFit(gset, design) # fit linear model

# set up contrasts of interest and recalculate model coefficients
cts <- c(paste(groups[1], "-", groups[2], sep = ""))
cont.matrix <- makeContrasts(
  contrasts = cts,
  levels = design
)
fit2 <- contrasts.fit(
  fit,
  cont.matrix
)

# compute statistics and table of top significant genes
fit2 <- eBayes(
  fit2,
  0.01
)
tT <- topTable(
  fit2,
  adjust = "fdr",
  sort.by = "B",
  number = 250
)

tT <- subset(
  tT,
  select = columns_for_tT
)

# write.table(
#   tT,
#   file = stdout(),
#   row.names = FALSE,
#   sep = "\t"
# )
produceDataTableWithButtons(tT)

gsubset <- gset[fData(gset)[["ID"]] %in% tT$ID, ]

set.seed(123)
Heatmap(
  exprs(gsubset),
  row_labels = fData(gsubset)$GENE_SYMBOL,
  top_annotation = HeatmapAnnotation(
    df = pData(gsubset)[
      ,
      c(
        "hybridization campaign:ch1",
        "tissue:ch1",
        "ucms:ch1",
        "flx:ch1"
      )
    ]
  ),
  name = "RNA signal",
  column_title = "Samples",
  row_title = "Genes"
)

# Visualize and quality control test results.
# Build histogram of P-values for all genes. Normal test
# assumption is that most genes are not differentially expressed.
tT2 <- topTable(
  fit2,
  adjust = "fdr",
  sort.by = "B",
  number = Inf
)
hist(
  tT2$adj.P.Val,
  col = "grey",
  border = "white",
  xlab = "P-adj",
  ylab = "Number of genes",
  main = "P-adj value distribution"
)

# summarize test results as "up", "down" or "not expressed"
dT <- decideTests(
  fit2,
  adjust.method = "fdr",
  p.value = 0.05
)

# Venn diagram of results
vennDiagram(
  dT,
  circle.col = palette()
)

# create Q-Q plot for t-statistic
t.good <- which(!is.na(fit2$F)) # filter out bad probes
qqt(fit2$t[t.good], fit2$df.total[t.good], main = "Moderated t statistic")

# volcano plot (log P-value vs log fold change)
colnames(fit2) # list contrast names
ct <- 1 # choose contrast of interest
volcanoplot(fit2,
  coef = ct, main = colnames(fit2)[ct], pch = 20,
  highlight = length(which(dT[, ct] != 0)), names = rep("+", nrow(fit2))
)

# MD plot (log fold change vs mean log expression)
# highlight statistically significant (p-adj < 0.05) probes
plotMD(fit2, column = ct, status = dT[, ct], legend = F, pch = 20, cex = 1)
abline(h = 0)

################################################################
# General expression data analysis
ex <- exprs(gset)

# box-and-whisker plot
ord <- order(gs) # order samples by group
palette(c(
  "#1B9E77", "#7570B3", "#E7298A", "#E6AB02", "#D95F02",
  "#66A61E", "#A6761D", "#B32424", "#B324B3", "#666666"
))
par(mar = c(7, 4, 2, 1))
title <- paste("GSE84185", "/", annotation(gset), sep = "")
boxplot(ex[, ord], boxwex = 0.6, notch = T, main = title, outline = FALSE, las = 2, col = gs[ord])
legend("topleft", groups, fill = palette(), bty = "n")

# expression value distribution
par(mar = c(4, 4, 2, 1))
title <- paste("GSE84185", "/", annotation(gset), " value distribution", sep = "")
plotDensities(ex, group = gs, main = title, legend = "topright")

# UMAP plot (dimensionality reduction)
ex <- na.omit(ex) # eliminate rows with NAs
ex <- ex[!duplicated(ex), ] # remove duplicates
ump <- umap(t(ex), n_neighbors = 7, random_state = 123)
par(mar = c(3, 3, 2, 6), xpd = TRUE)
plot(ump$layout, main = "UMAP plot, nbrs=7", xlab = "", ylab = "", col = gs, pch = 20, cex = 1.5)
legend("topright",
  inset = c(-0.15, 0), legend = levels(gs), pch = 20,
  col = 1:nlevels(gs), title = "Group", pt.cex = 1.5
)
library("maptools") # point labels without overlaps
pointLabel(ump$layout, labels = rownames(ump$layout), method = "SANN", cex = 0.6)

# mean-variance trend, helps to see if precision weights are needed
plotSA(fit2, main = "Mean variance trend, GSE84185")
```

```{r resultsMicroarrayGSE84185StressedFluoxetineVsStressedAloneInDentateGyrusDefaultSettingsToIntegrationOriginalSpecie}
# Rename Limma columns to match the ones from DESeq2
setnames(
  x = tT2,
  old = c(
    "AveExpr",
    "logFC",
    "P.Value",
    "adj.P.Val"
  ),
  new = c(
    "baseMean",
    "log2FoldChange",
    "pvalue",
    "padj"
  )
)

project <- "GSE84185"
comparison <- "StressedFluoxetineVsStressedAloneInDentateGyrus"
tT2$project <- project
tT2$comparison <- comparison
dea_res[["microarray_v2"]][[project]][[comparison]] <- tT2[
  !(
    tT2$Gene.name %like% "," |
    is.na(tT2$Gene.name) |
    duplicated(tT2$Gene.name)
  ),
  cols_to_for_integration_original_specie
]
```

```{r resultsMicroarrayGSE84185StressedFluoxetineVsStressedAloneInDentateGyrusDefaultSettingsToIntegration}
tT2 <- tT2[!tT2$ENSEMBL_ID %in% "", ]
# Probes are ordered by increasing p-value
# We keep only the one with the best pvalue for each gene
tT2 <- tT2[!duplicated(tT2$ENSEMBL_ID), ]

best_orth <- gorth(
  tT2$ENSEMBL_ID,
  source_organism = "mmusculus",
  target_organism = "hsapiens",
  mthreshold = 1,
  filter_na = FALSE
)

best_orth <- best_orth[
  ,
  !colnames(best_orth) %in% c(
    "input_number",
    "input_ensg",
    "ensg_number"
  )
]
setnames(
  best_orth,
  old = c(
    "ortholog_name",
    "ortholog_ensg"
  ),
  new = c(
    "hugo_symbol",
    "ensembl_gene_id"
  )
)

tT2 <- merge(
  tT2,
  best_orth,
  by.x = "ENSEMBL_ID",
  by.y = "input",
  all.x = TRUE,
  sort = FALSE
)

tT2 <- tT2[!tT2$hugo_symbol %in% "N/A", ]
dea_res[["microarray"]][[project]][[comparison]] <- tT2
```

```{r resultsMicroarrayGSE84185StressedFluoxetineVsStressedAloneInDentateGyrusDefaultSettingsHeatmapSettings}
fig_height <- 30
n_top_genes <- 200
gsubset <- gset[fData(gset)[["ID"]] %in% tT2[1:n_top_genes, ]$ID, ]
```

```{r resultsMicroarrayGSE84185StressedFluoxetineVsStressedAloneInDentateGyrusDefaultSettingsHeatmapRaw, dependson="resultsMicroarrayGSE84185StressedFluoxetineVsStressedAloneInDentateGyrusDefaultSettingsHeatmapSettings", fig.height=fig_height}
set.seed(123)
Heatmap(
  exprs(gsubset),
  row_labels = fData(gsubset)$GENE_SYMBOL,
  #row_labels = tT2[1:n_top_genes, ]$hugo_symbol,
  top_annotation = HeatmapAnnotation(
    df = pData(gsubset)[
      ,
      c(
        "hybridization campaign:ch1",
        "tissue:ch1",
        "ucms:ch1",
        "flx:ch1"
      )
    ]
  ),
  name = "RNA signal",
  column_title = "Samples",
  row_title = "Genes"
)
```

```{r resultsMicroarrayGSE84185StressedFluoxetineVsStressedAloneInDentateGyrusDefaultSettingsHeatmapZscore, dependson="resultsMicroarrayGSE84185StressedFluoxetineVsStressedAloneInDentateGyrusDefaultSettingsHeatmapSettings", fig.height=fig_height}
set.seed(123)
Heatmap(
  scaleRows(exprs(gsubset)),
  row_labels = fData(gsubset)$GENE_SYMBOL,
  #row_labels = tT2[1:n_top_genes, ]$hugo_symbol,
  top_annotation = HeatmapAnnotation(
    df = pData(gsubset)[
      ,
      c(
        "hybridization campaign:ch1",
        "tissue:ch1",
        "ucms:ch1",
        "flx:ch1"
      )
    ]
  ),
  name = "RNA signal",
  column_title = "Samples",
  row_title = "Genes"
)
```

This heatmap was used to check fold change sign is correct for this design:
As manually verified example, we can see FHL2 is overexpressed with Fluoxetine, which can be consistently seen on integration dotplots for the design `GSE84185-StressedFluoxetineVsStressedAloneInDentateGyrus`.
